---
title: "Unit 4 Assignment - Versioning"
author: "James Vasquez"
date: "May 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###### MSDS 6306: Doing Data Science
###### Thursday 6:30 - 8pm

1.   FiveThirtyEight data package
    a.   Install the 'fivethirtyeight' package: 'install.packages("fivethirtyeight")'
```{r install538, echo=FALSE}
library(fivethirtyeight)
library(plyr)
library(dplyr)
library(vcdExtra)
library(data.table)
library(urltools)
library(ggplot2)
```

1.   FiveThirtyEight data package
    b.   Assign 22nd data set to object called 'df'
```{r Assign df, echo=TRUE}
#Find the 18th data set in 538
DataSet18 <- tail(as.data.frame(head(vcdExtra::datasets("fivethirtyeight"), 22)),1)
DataSet18

#set the 18th data set to a data frame called df
df <- as.data.frame(college_recent_grads)
```

1.   FiveThirtyEight data package
    c.   Comment the URL for the 22nd data set
```{r df URL, echo=TRUE}
#Show URL for Data Set 18
DataSet18_address <- url_parse("https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/")
url_compose(DataSet18_address)
```

1.   FiveThirtyEight data package
    d.   Give the dimensions for the 22nd data set
```{r dim_df, echo=TRUE}
#Dimensions for df
dim(df)

#Column names listed vertically
df_ColNames <- matrix(colnames(df))
colnames(df_ColNames) <- c("col_name")
df_ColNames
```

2.   FiveThirtyEight Data Summary
    a.   List column names and number of columns in the data frame
```{r colnames_and_count_columns, echo=TRUE}
#List column names vertically
df_ColNames <- matrix(colnames(df))
colnames(df_ColNames) <- c("col_name")
df_ColNames

#Count number of columns in data frame
ncol(df)
```

2.   FiveThirtyEight Data Summary
    b.   Generate countof each unique 'major_category'
```{r major_count, echo=TRUE}
Major_Count <- as.data.frame(df %>% group_by(major_category) %>% summarize(count=n()))
Major_Count <- Major_Count[complete.cases(Major_Count),]
Major_Count <- Major_Count[order(Major_Count$count,decreasing=TRUE),c(1,2)]
Major_Count
```

2.   FiveThirtyEight Data Summary
    c.   Plot 'Major_Count' as horizontal bar plots with different color
```{r major_count_BarPlot, echo=TRUE}
Major_CountBarPlot <- ggplot(data=Major_Count, 
                             aes(x=reorder(major_category, count),
                             y=count,
                             fill=major_category)) + 
                             geom_bar(stat="identity") +
                             coord_flip() +
                             labs(x="Majors", 
                                  y="Number of occurence", 
                                  title = "Occurence counts of Majors")
#Show bar plot
Major_CountBarPlot
```

2.   FiveThirtyEight Data Summary
    d.   write df to CSV file
```{r write_CSV_file, echo=TRUE}
write.csv(df, "college_recent_grads.csv", row.names = FALSE)
```




###Session Information
```{r session, echo=FALSE}
sessionInfo()

```
