---
title: "Unit 3 Assignment - File Management"
author: "James Vasquez"
date: "May 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###### MSDS 6306: Doing Data Science
###### Thursday 6:30 - 8pm


1.   In R read in titanic.csv via  either read.table()  or read.csv(), assigning  it  to  df.    This  dataset  follows  the  passengers  aboard  the  Titanic,  including  their  fees  paid,  rooms  rented,  and  survivorship  status.

```{r Read Titanic Data, echo=TRUE}
library(plyr)

#Set working directory
TitanicDIR <- setwd("C:/Users/James/Documents/SMU/MSDS_6306_Doing_Data_Science/gitprojects/awesome-public-datasets/Datasets/titanic/")

#Read CSV file into R
TitanicData <- read.csv("titanic.csv", header=TRUE, sep=",")
MatrixTitanicData <- matrix(TitanicData)

#check data table class
class(TitanicData)
class(MatrixTitanicData)
```

2. Frequencies and graphs output
  i) count  of  females  and  males  
  i) Plot  the  frequency  of  females  and  males.    
    A.   Be  sure  to  give  an  accurate  title  and  label  the  axes

```{r Gender Frequency, echo=TRUE}
#Get frequency table of the Sex column
SexFreq <- count(TitanicData, 'Sex')
countsSEX <- table(TitanicData$Sex)

#show results
SexFreq
countsSEX

#Make bar chart
barplot(countsSEX, 
        main="Gender Frequency Count", 
        xlab="Gender",
        ylab = "Counts",
        col = c("darkblue"))
```

3. Please  use  one  apply  function to output (values should be real numbers): 
i. Mean  of  Age  
i. Mean of fare 
i. Mean of survival

```{r Means, echo=TRUE}
#Extract columns needed to find means
AgeVector <- as.vector(TitanicData$Age)
FareVector <- as.vector(TitanicData$Fare)
SurvivalVector <- as.vector(TitanicData$Survived)

#Combine vectors into matrix
MATRIXAgeFareSurvived <- cbind(AgeVector, FareVector, SurvivalVector)

#output means of Age, Fare, Survival. Output should be real number
MeanAgeFareSurvived <- apply(MATRIXAgeFareSurvived, 2, mean, na.rm=TRUE)

#shows output is numeric
class(MeanAgeFareSurvived)

#Show Means
MeanAgeFareSurvived
```

4. Create function to do the following
  i) Create objects for 
    A.   Median age 
    A.   Min duration sleep
    A.   Max duration sleep 
    A.   Mean RSES
    A.   Standard deviation RSES
  i) Create data frame called 'Report'
    A.   Divide results by 5
  i) Columns names should be
    A.   MedianAge
    A.   SelfEsteem
    A.   SE_SD
    A.   DurationRange
  i) Round the report to 2 sig figs

```{r Function Building, echo=TRUE}
#Set sleep data working directory
SleepDIR <- setwd("C:/Users/James/Documents/SMU/MSDS_6306_Doing_Data_Science/HW_Data/Unit3_File_Mgmt/")

#Read CSV file into R
SleepData <- read.csv("sleep_data_01.csv", header=TRUE, sep=",")

HW3FileMgmt <- function(x) {

  #create vectors
  AgeVector <- as.vector(SleepData$Age)
  DurationVector <- as.vector(SleepData$Duration)
  RSESVector <- as.vector(SleepData$RSES)
  
  #Create metric objects
  MedianAge <- mean(AgeVector, na.rm=TRUE)
  DurationMin <- min(DurationVector, na.rm=TRUE)
  DurationMax <- max(DurationVector, na.rm=TRUE)
  SelfEsteem <- mean(RSESVector, na.rm=TRUE)
  SE_SD <- sd(RSESVector, na.rm=TRUE)
  
  #Create data frame and divide results by 5, limit to 2 sig figs
  Report <- round(data.frame(cbind(MedianAge, DurationMin, DurationMax, SelfEsteem, SE_SD))/5, digits = 2)
  
  return(Report)
}
```

5. Function Results!!!!
```{r Function Results, echo=TRUE}
#function(data frame name)
HW3FileMgmt(SleepData)
```

###Session Information
```{r session, echo=FALSE}
sessionInfo()

```
