---
title: "Unit 5 Assignment - Tidy Data"
author: "James Vasquez"
date: "June 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###### MSDS 6306: Doing Data Science
###### Thursday 6:30 - 8pm

1.   Data Munging
    a.   Import 'yob2016.txt' and assign column names and assign as 'df'
```{r Import_yobData, echo=TRUE}
library(data.table)
library(sqldf)
library(RSQLite)
library(DBI)
library(gsubfn)
library(proto)

#Read in the yob2016.txt data set
df <- read.table("yob2016.txt", header = FALSE, sep = ";", strip.white = TRUE)
colnames(df)[colnames(df)=='V1'] <- 'Name'
colnames(df)[colnames(df)=='V2'] <- 'Gender'
colnames(df)[colnames(df)=='V3'] <- 'Count_of_Name'
head(df,2)
```

1.   Data Munging
    b.   Display summary and structure of 'df'
```{r DisplayStructure, echo=TRUE}
#Get dimensions of df
dim(df)
#show class
class(df)
#show summary
summary(df)
#Show Structure of df
str(df)
```

1.   Data Munging
    c.   Find a misspelled name with the last letters '%yyy'
```{r find_misspelled, echo=TRUE}
#find misspelled names

sqldf("Select * 
      FROM df 
      WHERE Name like '%yyy';")

```

1.   Data Munging
    d.   Remove the misspelled names from the data and assign the table as 'y2016'
```{r munging, echo=TRUE}
y2016 <- df[!(df$Name=="Fionayyy"),]
#Show dimensions before row is deleted
dim(df)
#Show dimensions after row is deleted
dim(y2016)
```

2.   Data Merging
    a.   Import 'yob2015.txt' and assign column names and assign as 'y2015'
```{r import_yob2015, echo=TRUE}
y2015 <- read.table("yob2015.txt", header = FALSE, sep = ",", strip.white = TRUE)
colnames(y2015)[colnames(y2015)=='V1'] <- 'Name'
colnames(y2015)[colnames(y2015)=='V2'] <- 'Gender'
colnames(y2015)[colnames(y2015)=='V3'] <- 'Count_of_Name'
head(y2015, 2)
```
```{r}

```

2.   Data Merging
    b.   Display last 10 rows of 'y2015' and describe what you see
```{r Display_last_10, echo=TRUE}
tail(y2015,10)

#They are all males and have 5 obs each
```

2.   Data Merging
    c.   Merge y2016 and y2015 and assign to 'final', there should be no NA's
```{r Merge_df, echo=TRUE}
#merge data frames on name and gender
final <- merge(y2015, y2016, by.x=c("Name","Gender"), by.y=c("Name","Gender"), all=FALSE)

#rename column names
colnames(final)[colnames(final)=='Count_of_Name.x'] <- 'Count_of_Name_2015'
colnames(final)[colnames(final)=='Count_of_Name.y'] <- 'Count_of_Name_2016'

head(final, 2)
```

3.   Data Summary
    a.   Create new column 'Total' that adds both the 2015 and 2016 counts and SUM this column
    b.   Sort 'final' by Total and show top 10 names
```{r TotalSum, echo=TRUE}
#Add the counts and creates Total column
#sorts by Total in descending order
final <- sqldf("Select 
          Name,
          Gender,
          Count_of_Name_2015,
          Count_of_Name_2016,
          Count_of_Name_2015+Count_of_Name_2016 as Total
      FROM final
      ORDER by Total desc;")

#Report sum total names
sqldf("Select SUM(Count_of_Name_2015+Count_of_Name_2016) as TotalCountofNames FROM final;")
head(final, 10)
```

3.   Data Summary
    c.   Show top 10 female names
```{r Top10_female_names, echo=TRUE}
final_Top10_female_names <- sqldf("Select 
                        Name,
                        Count_of_Name_2015+Count_of_Name_2016 as Total
                        FROM final
                        WHERE Gender = 'F'
                        ORDER by Total desc limit 10;")
final_Top10_female_names
```

3.   Data Summary
    d.   Write 'final' as *.CSV file
```{r writedata, echo=TRUE}
#write.csv(final_Top10_female_names, file="Top_10_Popular_Female_Names.csv", row.names = FALSE)
```

4.   Upload to github repository 
    a.   [James Vasquez GitHub Link](https://github.com/jvasquezDC/SMU-MSDS6306-HW.git)

###Session Information
```{r session, echo=FALSE}
sessionInfo()

```
